---
title: "Data 609 Module 7 HW"
author: "Bryan Persaud"
date: "5/16/2021"
output:
  pdf_document: default
  html_document: default
---

# Ex.1

Use the svm() algorithm of the e1071 package to carry out the support vector machine for the PlantGrowth data set. Then, discuss the number of support vectors/samples. [Install the e1071 package in R if needed.]

Solution:

```{r}
library(e1071)
plant_svm <- svm(group ~ ., data = PlantGrowth)
summary(plant_svm)
```

When using the svm() algorithm on the PlantGrowth data set it shows that there are 29 support vectors.

# Ex.2

Do a similar SVM analysis as that in the previous question using the iris data set. Discuss the number of support vectors/samples.

Solution:

```{r}
iris_svm <- svm(Species ~ ., data = iris)
summary(iris_svm)
```

When using the svm() algorithm on the iris data set it shows that there are 51 support vectors.

# Ex.3

Use the iris data set (or any other data set) to select 80% of the samples for the training svm(), then use the rest 20% for validation. Discuss your results.

Solution:

```{r}
library(caret)
train <- createDataPartition(iris$Species, p = 0.80, list = FALSE)
validation <- iris[-train]
iris_dataset <- data.frame(iris[train])
svm(train ~ ., data = iris_dataset)
```

Selecting 80% of the samples to train svm() gives us 120 support vectors.