---
title: "Data 609 Module 3 HW"
author: "Bryan Persaud"
date: "3/14/2021"
output:
  pdf_document: default
  html_document: default
---

# Ex.1

Write down Newton's formula for finding the minimum of $f(x) =  (3x^{4} -  4x^{3}) / 12$ in the range [-10, 10]. Then, implement it in R.

Solution:

First find the first two derivatives of f(x) 

$f'(x) =  \frac{ 12x^{3} -  12x^{2} }{12} =  x^{2}(x - 1) = x^{3} - x^{2}$

$f''(x) = 3x^{2} - 2x$

Then we know Newton's formula to be $x_{k  + 1} =  x_{k} -  \frac{x_k^{3} - x_k^{2}}{3x_k^{2} - 2x_k}$

```{r}
f <- function(x) {
 x - (x^3 - x^2) / (3 * x^2 - 2 * x)
}
f1 <- function(X_0) {
  values <- c()
  for(i in 1:10) {
    if(i == 1){
      values[[i]] <- f(X_0)
    }
    else{
      values[[i]] <- f(values[[i - 1]])
    }
  }
  print(values)
}
f1(2)
```

```{r}
f1(5)
```

We see that the minimum is 1 when multiple starting values are used.

# Ex.2

Explore optimize() in R and try to solve the previous problem.

Solution:

```{r}
f <- function(x) {
  (3*x^4 - 4*x^3) / 12
}
xmin <- optimize(f, interval = c(-10,10), tol = 0.0001)
xmin
```

We see that the minimum is close to 1 with a value for f(x) at -0.083.

# Ex.3

Use any optimization algorithm to find the minimum of $f(x,y) =  (x-1)^{2} + 100 (y -  x^{2}) ^{2}$ in the domain $-10  \leq x,y  \leq 10.$ Discuss any issues concerning the optimization process.

Solution:

Newton method for a higher dimension will be used to answer the question. 

$x_{t + 1} =  x_{t} -  H^{-1}\bigtriangledown f(x,y)$

$x_{0} = \begin{bmatrix}x  \\y \end{bmatrix} =\begin{bmatrix}0  \\0 \end{bmatrix}$ Starting value

$\bigtriangledown f(x,y) =  \begin{bmatrix}(x-1)^{2}  \\100 (y -  x^{2}) ^{2} \end{bmatrix}$

find the Hessian matrix by finding second derivative and plugging in starting values

$f_{xx} = 1200 x^{2} + 2 - 400y$

$f_{xy} = -400x$

$f_{yy} = 200$

$f_{yx} = -400x$

$H = \bigtriangledown^2 f(x,y) =  \begin{bmatrix}2 & -0  \\-0 & 200 \end{bmatrix}$ 

$H^{-1}  = \begin{bmatrix}0.5 & 0  \\0 & 0.005 \end{bmatrix}$ 

With t = 0 $x_{1} =  x_{0} -  H^{-1}\bigtriangledown f(x,y)$

$x_{1} =  \begin{bmatrix}0  \\0 \end{bmatrix} - \begin{bmatrix}0.5 & 0  \\0 & 0.005 \end{bmatrix}\begin{bmatrix}1  \\0 \end{bmatrix}$

$x_{1} =  \begin{bmatrix}0  \\0 \end{bmatrix} - \begin{bmatrix}0.5  \\0 \end{bmatrix}$

$x_{1} =  \begin{bmatrix}-0.5  \\0 \end{bmatrix}$

From here f(x,y) will converge into a single point as you go through more values of t. The issues I faced was optimizing for a multivariate function. I ended up choosing this optimization algorithm as it was the only one I knew how to use.

# Ex.4

Explore the optimr package for R and try to solve the previous problem.

Solution:

```{r}
library(optimr)
```

```{r}
fn <- function(x, y){
  (x - 1)^2 + 100 * (y - x^2)^2
}
optimr(c(-10,10), fn, method = "L-BFGS-B")
```