---
title: "Data 605 HW 13"
author: "Bryan Persaud"
date: "5/3/2020"
output: html_document
---

# 1

Use integration by substitution to solve the integral below.

$\int { 4{ e }^{ -7x }dx }$

Solution:

Let u = -7x, then du = -7dx. In terms of dx, dx = $-\frac { 1 }{ 7 } du$

Rewrite integral in terms of u.

$\int { 4{ e }^{ u }(-\frac { 1 }{ 7 }) du }$ Take out constants.

$-\frac { 4 }{ 7 }\int { { e }^{ u } du }$ Integrate.

$-\frac { 4 }{ 7 }({ e }^{ u } + C)$ Substitute back in terms of x.

$-\frac { 4 }{ 7 }({ e }^{ -7x } + C)$ For some constant C.

# 2

Biologists are treating a pond contaminated with bacteria. The level of contamination is changing at a rate of $\frac { dN }{ dt } =-\frac { 3150 }{ { t }^{ 4 } } -220$ bacteria per cubic centimeter per day, where t is the number of days since treatment began. Find a function N( t ) to estimate the level of contamination if the level after 1 day was 6530 bacteria per cubic centimeter.

Solution:

$\int { -\frac { 3150 }{ { t }^{ 4 } } -220 }dt$ Split integral.

$\int { -\frac { 3150 }{ { t }^{ 4 } }dt -\int { 220 }dt  }$ Take out constants for both integrals

$-3150\int { \frac { 1 }{ { t }^{ 4 }dt } -\int { 220 }dt  }$ Integrate.

$-3150(-\frac { 1 }{ 3{ t }^{ 3 } } )-220t$ Simplify.

$\frac { 1050 }{ { t }^{ 3 } } -220t$ Write as function N(t).

$N(t)=\frac { 1050 }{ { t }^{ 3 } } -220t+C$ For some constant C.

Solve for N(t) = 6530 where t = 1.

$\frac { 1050 }{ { t }^{ 3 } } -220t+C=6530$

$\frac { 1050 }{ { 1 }^{ 3 } } -220(1)+C=6530$

$1050-220+C=6530$

$830+C=6530$

$C=5700$

Function N(t)= $\frac { 1050 }{ { t }^{ 3 } } -220t+5700$

# 3

Find the total area of the red rectangles in the figure below, where the equation of the line is f ( x ) = 2x - 9.

Solution:

From looking at the graph, it looks like there are 4 red rectangles that start at 4.5 and end at 8.5.

```{r}
function_3 <- function(x)
{
  2 * x - 9
}
integrate(function_3, lower = 4.5, upper = 8.5)
```

The total area of the red rectangles is 16.

# 4

Find the area of the region bounded by the graphs of the given equations.

$y={ x }^{ 2 }-2x-2,y=x+2$

Enter your answer below.

Solution:

Find the intersection of the given equations

${ x }^{ 2 }-2x-2=x+2$ Set equal to 0.

${ x }^{ 2 }-3x-4=0$ Factor.

$(x-4)(x+1)=0$ Solve for x.

$x=4,x=-1$

Subtract the two equations.

$x+2-({ x }^{ 2 }-2x-2)$ Simplify.

$x+2-{ x }^{ 2 }+2x+2$ Combine like terms.

$-{ x }^{ 2 }+3x+4$

Integrate this equation using intersection points as interval.

```{r}
function_4 <- function(x)
{
  -x ^ 2 + 3 * x + 4
}
integrate(function_4, lower = -1, upper = 4)
```

The area of the region bounded by the graphs of the given equations is 20.83333.

# 5

A beauty supply store expects to sell 110 flat irons during the next year. It costs $3.75 to store one flat iron for one year. There is a fixed cost of $8.25 for each order. Find the lot size and the number of orders per year that will minimize inventory costs.

Solution:

The lot size will be 110 / the number of orders per year. Let us have n = number of orders per year and lot size = l. Then l = 110 / n. Our cost function will look like $8.25n\quad +\quad 3.75(\frac { l }{ 2 } )$ This function simplified is

$8.25n + 3.75(\frac { l }{ 2 } )$ Plug in l.

$8.25n + 3.75(\frac { \frac { 110 }{ n }  }{ 2 } )$ Simplify fraction.

$8.25n+(\frac { 110*3.75 }{ 2n } )$ Cost function is equal to.

$8.25n-(\frac { 206.25 }{ n } )$

Take the derivative of the cost function.

$8.25n-(\frac { 206.25 }{ n } )$ derivative =

$8.25-(\frac { 206.25 }{ { n }^{ 2 } } )$

Solve for n by setting derivative of cost function equal to 0.

$8.25-(\frac { 206.25 }{ { n }^{ 2 } } )=0$ Move 8.25 to other side.

$-(\frac { 206.25 }{ { n }^{ 2 } } )=-8.25$ Cross multiply.

$-8.25{ n }^{ 2 }=-206.25$ Divide -8.25.

${ n }^{ 2 }=\frac { -206.25 }{ -8.25 }$ Simplify.

${ n }^{ 2 }=25$ Solve for n.

$n=5$

Now that we know n we can solve for l.

$l=\frac { 110 }{ n }$ Plug in 5 for n.

$l=\frac { 110 }{ 5 }$ Simplify.

$l=22$

The lot size and the number of orders per year that will minimize inventory costs are a lot size of 22 and the number of orders per year equal to 5. 

# 6

Use integration by parts to solve the integral below.

$\int { ln(9x)·{ x }^{ 6 }dx }$

Solution:

Let u = ln(9x), then du = $\frac { 1 }{ x } dx$. Let v = $\frac { 1 }{ 7 } { x }^{ 7 }$, then dv = ${ x }^{ 6 }$.

Using integration by parts, we get u * v - the integral of v * du.

$\ln { (9x)(\frac { 1 }{ 7 } { x }^{ 7 })-\int { \frac { 1 }{ 7 } { x }^{ 7 }(\frac { 1 }{ x } dx) }  }$ Simplify.

$\frac { { x }^{ 7 }ln(9x) }{ 7 } - \int { \frac { { x }^{ 6 } }{ 7 } dx }$ Integrate.

$\frac { { x }^{ 7 }ln(9x) }{ 7 } - (\frac { { x }^{ 7 } }{ 49 } +C)$ For some constant C.

# 7

Determine whether f ( x ) is a probability density function on the interval [1, e ^ 6] . If not, determine the value of the definite integral.

$f\left( x \right) =\frac { 1 }{ 6x } $

Solution:

```{r}
function_7 <- function(x)
{
  1 / (6 * x)
}
integrate(function_7, lower = 1, upper = exp(6))
```

f(x) is a probability density function on the interval [1, e ^ 6] since the integral of f(x) on  the interval is equal to 1.