---
title: "Data 605 HW 2"
author: "Bryan Persaud"
date: "2/9/2020"
output: html_document
---

# Problem Set 1

(1) Show that ${ A }^{ T }A \neq A{ A }^{ T }$ in general. (Proof and demonstration.)

Solution:

Proof:

Suppose we have A that is a m x n matrix. The transpose of A, ${ A }^{ T }$, would be a n x m matrix. Using the definition of matrix multiplication, we can multiply ${ A }^{ T }$ by A which would result in a n x n matrix; a square matrix. If we multiply A by ${ A }^{ T }$ this would result in a m x m matrix; also square matrix. Since you end up with two different square matrices, one that is n x n and the other that is m x m, ${ A }^{ T } A \neq A{ A }^{ T }$.

Demonstration:

Let A be a 2 x 3 matrix. Then the transpose of A is ${ A }^{ T }$, a 3 x 2 matrix. Then ${ A }^{ T }A$ would be a 3 x 3 matrix and  $A{ A }^{ T }$ would be a 2 x 2 matrix.

A = $\begin{bmatrix} 1 & 2\quad 3 \\ 3 & 2\quad 1 \end{bmatrix}$

${ A }^{ T }$ = $\begin{bmatrix} 1 & 3 \\ 2 & 2 \\ 3 & 1 \end{bmatrix}$

${ A }^{ T }A$ = $\begin{bmatrix} 1 & 3 \\ 2 & 2 \\ 3 & 1 \end{bmatrix}$ * $\begin{bmatrix} 1 & 2\quad 3 \\ 3 & 2\quad 1 \end{bmatrix}$

= $\begin{bmatrix} 10 & 8 & 6 \\ 8 & 8 & 8 \\ 6 & 8 & 10 \end{bmatrix}$

$A{ A }^{ T }$ = $\begin{bmatrix} 1 & 2\quad 3 \\ 3 & 2\quad 1 \end{bmatrix}$ * $\begin{bmatrix} 1 & 3 \\ 2 & 2 \\ 3 & 1 \end{bmatrix}$

= $\begin{bmatrix} 14 & 10 \\ 10 & 14 \end{bmatrix}$ 

As you can see above, ${ A }^{ T }A \neq A{ A }^{ T }$.

(2) For a special type of square matrix A, we get ${ A }^{ T }A = A{ A }^{ T }$. Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).

Solution:

The conditions that could make ${ A }^{ T }A = A{ A }^{ T }$ for a special type of square matrix A is if A and ${ A }^{ T }$ are equal to each other. Since A and ${ A }^{ T }$ are equal to each other, this would mean the special type of square matrix would be a symmetric matrix. For example the Identity matrix is a special type of square matrix that shows this since it is a square matrix whose transpose would be equal to itself. In conclusion, if a square matrix A is a symmetric matrix, meaning that A and ${ A }^{ T }$ are equal to each other, then ${ A }^{ T }A = A{ A }^{ T }$. 

# Problem Set 2

```{r}
Matrix_Factorization <- function(A) {
  x <- dim(A)[1]
  D <- matrix(diag(x), nrow = x, ncol = x)
  L <- D
  U <- A
  for (b in 1:(x - 1)) {
    for (a in (b + 1):x) {
      L[a,b] <- U[a,b] / U[b,b]
      U[a,] <- U[a,] - U[b,] * L[a,b]
    }
  }
  diag(D) <- diag(U)
  for (c in 1:x) {
    U[c,] <- U[c,] / U[c,c]
  }
  LDU = list("Lower" = L, "Diagonal" = D, "Upper" = U)
  return(LDU)
}
```

Test to see if function works:

```{r}
A <- matrix(c(1, 3, 4, 2, 3, 1, 4, 1, 3), nrow = 3, byrow = TRUE)
Matrix_Factorization(A)
```