---
title: "Discussion 9"
author: "Bryan Persaud"
date: "3/24/2020"
output: html_document
---

# Chapter 9 Section 9.2 Exercise 5

A die is thrown until the first time the total sum of the face values of the die is 700 or greater. Estimate the probability that, for this to happen,

(a) more than 210 tosses are required.
(b) less than 190 tosses are required.
(c) between 180 and 210 tosses, inclusive, are required.

Solution:

First lets calculate the total sum of the face values of the die. 

${ S }_{ n }={ X }_{ 1 }+{ X }_{ 2 }+...+{ X }_{ n }$

The expected value and variance would be:

$E(X)=\frac { 7 }{ 2 }$ and $E({ S }_{ n })=n*\frac { 7 }{ 2 }$

$V(X)=\frac { 35 }{ 12 } $ and $V({ S }_{ n })=n*\frac { 35 }{ 12 }$ which means that $σ({ S }_{ n })=\sqrt { n*\frac { 35 }{ 12 }  }$

(a)

$P({ S }_{ 210 }<700)$

$E({ S }_{ 210 })=210*\frac { 7 }{ 2 }=735$
 
$σ({ S }_{ 210 })=\sqrt { 210*\frac { 35 }{ 12 }  }=24.75$
 
```{r}
round((700 - 735) / 24.75, 4)
```
 
```{r}
round(pnorm(-1.4141), 4)
```

$P({ S }_{ 210 }<700)=0.0787$

The probability that the total sum of the face values of the die requires more than 210 tosses is 0.0787.

(b)

$P({ S }_{ 190 }\ge 700)$

$E({ S }_{ 190 })=190*\frac { 7 }{ 2 }=665$
 
$σ({ S }_{ 190 })=\sqrt { 190*\frac { 35 }{ 12 }  }=23.54$

```{r}
round((700 - 665) / 23.54, 4)
```

```{r}
round(1 - pnorm(1.4868), 4)
```

$P({ S }_{ 190 }\ge 700)$

The probability that the total sum of the face values of the die requires less than 190 tosses is 0.0685.

(c)

$P({ S }_{ 180 }<700, P({ S }_{ 210 }\ge 700)$ =

$P({ S }_{ 180 }<700)-P({ S }_{ 180 }<700, P({ S }_{ 210 }\ge 700)$ =

$P({ S }_{ 180 }<700)-P({ S }_{ 210 }\ge 700)$

$P({ S }_{ 180 }<700)$

$E({ S }_{ 180 })=180*\frac { 7 }{ 2 }=630$
 
$σ({ S }_{ 180 })=\sqrt { 180*\frac { 35 }{ 12 }  }=22.91$

```{r}
pnorm(22.91)
```

```{r}
1 - 0.0757
```

The probability that the total sum of the face values of the die requires between 180 and 210 tosses is 0.0685.