---
title: "Discussion 13"
author: "Bryan Persaud"
date: "4/27/2020"
output: html_document
---

# Chapter 7 Section 7.1 Exercise 13

In Exercises 13 – 20, find the total area enclosed by the functions f and g.

13. $f(x) = 2x^2 + 5x − 3, g(x) = x^2 + 4x − 1$

Solution:

Let's use R to create two function, f(x) and g(x).

```{r}
function_f <- function(x)
{
  2 * (x ^ 2) + 5 * x - 3
}
function_g <- function(x)
{
  (x ^ 2) + 4 * x - 1
}
```

We need to find the intersection of the two functions. To do this we set f(x) = g(x).

$2x^2 + 5x - 3 = x^2 + 4x - 1$ = Subtract to set equal to 0.

$x^2 + x - 2 = 0$ Solve using the quadratic formula where a = 1, b = 1, and c = -2

I'll skip over the steps for the quadratic formula, but we end up with x = -2 and x = 1. So we have two intersection points, -2 and 1.

Now that we know the intersection points, we integrate f(x) - g(x), where our interval is made up of the two intersection points we found. We know that f(x) - g(x) is $x^2 + x - 2$.

```{r}
function_fg <- function(x)
{
  (x ^ 2) + x - 2
}
integrate(function_fg, lower = -2, upper = 1)
```

The total area enclosed by the functions f and g is -4.5. We can't have a negative area, so we take the absolute value of the total area. So we get 4.5 as the total area enclosed by the functions f and g.