---
title: "Data 607 Project 1"
author: "Bryan Persaud"
date: "9/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(stringr)
```



```{r}
# Get tournament data from tournament.txt that I uploaded to github. 
theURL <- "https://raw.githubusercontent.com/bpersaud104/Data607/master/tournamentinfo.txt"
Chess <- read.table(file = theURL, header = FALSE, sep = ",")
# Remove the header from data.
Chess <- Chess[-c(1:4),]
head(Chess)
```

```{r}
class(Chess)
```

This needs to change to character.

```{r}
Chess_char <- as.character(Chess)
class(Chess_char)
```


```{r}
# Seperate data by players
Players <- Chess_char[seq(from = 1, to = length(Chess_char), by = 3)]
Players
```

```{r}
# Seperate data by player's ratings
Ratings <- Chess_char[seq(from = 2, to = length(Chess_char), by = 3)]
Ratings
```

# Extract Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and  Pre Chess Rating of Opponents from seperated data.

```{r}
Player_Name <- str_trim(str_extract(Players, "(\\w+\\s){2,3}"))
Player_State <- str_extract(Ratings, "\\w\\w")
Total_Points <- as.numeric(str_extract(Players, "\\d.\\d"))
Players_Pre_Rating <- as.integer(str_extract(str_extract(Ratings, "\\D\\d{3,4}\\D"), "\\d+"))
Pre_Rating_Opponent <- str_extract_all(str_extract_all(Players, "\\d+\\|"), "\\d+")
Player_ID <- as.integer(str_extract(Players, "\\d+"))
```
str_extract caused NAs to be generated. I used str_extract twice to get rid of all the NAs. I used str_extract_all twice to remove "|" between each number. Kept getting an "argument is not an atomic vector; coercing", ignored message since code was still working.

# Create a new data frame to calculate average pre chess ratings of opponents.

```{r}; 
Played_Rounds <- sapply(Pre_Rating_Opponent, length)
Num_Rounds <- seq_len(max(Played_Rounds)) # Rounds 1 to 7
Results <- t(sapply(Pre_Rating_Opponent, "[", i = Num_Rounds))
Chess_data <- as.data.frame(Results)
Chess_data
```

I need to replace each entry in the data frame with the player's rating.

```{r}
Entry <- data.frame(Player_ID, Players_Pre_Rating)
Chess_Tournament <- Chess_data
Chess_Tournament[] <- Entry$Players_Pre_Rating[match(unlist(Chess_Tournament), Entry$Player_ID)]
Chess_Tournament
```


```{r}
Chess_Tournament$Avg <- round(apply(Chess_Tournament, 1, mean, na.rm = TRUE))
Chess_Tournament
```

# Combine everything into a single data frame

```{r}
Chess_Final <- data.frame(Player_Name, Player_State, Total_Points, Players_Pre_Rating, Chess_Tournament$Avg)
colnames(Chess_Final) <- c("Player's Name", "Player's State", "Total Number of Points", "Player's Pr-Rating", "Average Pre Chess Rating of Opponents")
head(Chess_Final, 64)
```

# Generate a .csv file

```{r}
write.csv(Chess_Final, file = "Data 607 Project 1.csv")
```

