---
output:
  html_document: default
  pdf_document: default
---
# Bryan Persaud
---
title: "Data 607 Week 3 Assignment"
author: "Bryan Persaud"
date: "9/15/2019"
output:
  pdf_document: default
  html_document: default
---

# 3.
```{r}
library(stringr)
```

```{r}
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
```
  # 1.
```{r}
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
```
# Need to fix "Simpson, Homer" and "Burns, C. Montgomery to conform to standard first_name last_name
```{r}
# Seperate the ones that are good into a new vector
good_names <- c(name[1], name[3], name[4], name[6])
good_names
```
```{r}
# Get Burns and Homer fixed by putting them in their own vector
Burns_name <- c(name[2])
Homer_name <- c(name[5])
Fix_Burns <- unlist(str_split(Burns_name, ",")) 
Fix_Burns
Fix_Homer <- unlist(str_split(Homer_name, ","))
Fix_Homer
```
```{r}
# Get rid of white spaces using str_trim
Fix_Burns <- str_trim(Fix_Burns)
Fix_Burns
Fix_Homer <- str_trim(Fix_Homer)
Fix_Homer
```
```{r}
# Switch to first_name last_name
Fix_Burns <- str_c(Fix_Burns[2], Fix_Burns[1], sep = " ")
Fix_Burns
Fix_Homer <- str_c(Fix_Homer[2], Fix_Homer[1], sep = " ")
Fix_Homer
```
```{r}
# Combine the ones that were already good to the ones that needed to be fixed into a new vector
The_Simpsons <- c(good_names, Fix_Burns, Fix_Homer)
The_Simpsons
```

  # 2.
```{r}
The_Simpsons
```

```{r}
Check_for_Title <- str_detect(The_Simpsons, "[[:alpha:]]{2,}\\.")
Check_for_Title
```
# Two Simpsons characters from this vector have titles, Rev. Timothy Lovejoy and Dr. Julius Hibbert

  # 3.
```{r}
The_Simpsons
```

```{r}
Check_for_Secondary_Name <- str_detect(The_Simpsons, "[A-Z]\\.{1}")
Check_for_Secondary_Name
```
# There is only one Simpsons character with a second name, C. Montgomery Burns. The C. stands for Charles 

# 4.

  # 1. 
```{r}
# [0-9]+\\$ shows a numerical value from numbers 0 to 9 that are followed by a $.

# Example to show the ones that match the regular expression.
Example1_vector <- c('123456789$', '123456789', 'Hello World', '1$', '789$')
Example1 <- str_extract(Example1_vector, "[0-9]+\\$")
Example1
```

  # 2.
```{r}
# \\b[a-z]{1,4}\\b shows letters from a to z that are lowercased and range from length 1 to 4.

# Example to show the ones that match the regular expression. 
Example2_vector <- c('DATA', 'data', '0000', 'Example two', 'Simpsons')
Example2 <- str_extract(Example2_vector, "\\b[a-z]{1,4}\\b")
Example2
```

  # 3.
```{r}
# .*?\\.txt$ shows any string of characters that end in .txt.

# Example to show the ones that match the regular expression.
Example3_vector <- c('example14.txt', 'intro to r.txt', 'intro to r.pdf', '123456789.txt', 'vector')
Example3 <- str_extract(Example3_vector, ".*?\\.txt$")
Example3
```

  # 4.
```{r}
# \\d{2}/\\d{2}/\\d{4} shows two numerical values followed by a / followed by two more numerical values followed by another / followed by four more numerical values. It basically shows a date format of ##/##/####.

# Example to show the ones that match the regular expression.
Example4_vector <- c('12/21/2012', 'ab/cd/efgh', '9/15/2019', '09/15/2019', '07/04/1776')
Example4 <- str_extract(Example4_vector, "\\d{2}/\\d{2}/\\d{4}")
Example4
```

  # 5.
```{r}
# <(.+?)?.+?</\\1> shows any string of characters inside of <> followed by more string of characters followed by another string of characters inside of <> but with a / inside of <>.

# Example to show the ones that match the regular expression.
Example5_vector <- c('<hw3> Character manipulation </hw3>', '<1234> MySQL<1234>', '<another> Example <another>', '<title> How to code using R </title>')
Example5 <- str_extract(Example5_vector, "<(.+?)?.+?</\\1>")
Example5
```

# 9.
```{r}
Secret_Message <- c("clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr")
Secret_Message
```
# Some character are uppercased. I'm going to pull out those characters to see if it will reveal the secret message.
```{r}
str_extract_all(Secret_Message, "[[:upper:]]")
```
# Hidden message is revealed!
