---
title: "Data 607 Assignment 9"
author: "Bryan Persaud"
date: "10/27/2019"
output: html_document
---

# I am choosing to use the New York Time's movie reviews API

```{r}
library(httr)
library(jsonlite)
library(knitr)
library(DT)
library(tidyr)
library(dplyr)
library(stringr)
```



```{r}
URL <- "https://api.nytimes.com/svc/movies/v2/reviews/all.json"
Movie_reviews <- GET(URL, query = list("api-key" = "POyKM5jA9Ge2v14RA5YGz49jkJim3gj9"))
Movie_content <- content(Movie_reviews, "text")
Movie_reviews
```

# Store JSON data into a dataframe

```{r}
Movie_data <- fromJSON(Movie_content)
Movie_df <- as.data.frame(Movie_data$results)
Movie_df
```

The dataframe seems to contain 20 cases, each case being a different movie that was recently released. Each case has information on the movie, such as the name of the movie, the release date, the rating of the movie, and  a link containing information on the movie. Also there is information on the movie critic reviewing the movie, such as their name, a title for their review, and a short description of the review.

# Some data analysis

Let's see how many R-rated movies were reviewed.

```{r}
Movie_df %>%
filter(str_detect(mpaa_rating, "R"))
```

Looks like out of the 20 movies that were reviewed, 5 of them were R rated movies.